### 2023-11-13 lmcs/priority-cp.tex
restructuring before reworking
Update lmcs.cls

\subsection{Revisiting Priority CP}
\begingroup
\usingnamespace{pcp}
\textbf{Types} ($\ty{A}$, $\ty{B}$) in PCP correspond to linear logic connectives annotated with priorities $\cs{o}\in\mathbb{N}$. Typing environments, duality, and the priority function $\pr(\cdot)$ are defined as expected.
% (see~\cref{sec:pcp-environments,sec:pcp-duality,sec:pcp-priorities}).
\[
  \begin{array}{lcl}
    \ty{A}, \ty{B}
    & \Coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
      \sep        \ty{\typarr[\cs{o}]{A}{B}}
      \sep        \ty{\tyone[\cs{o}]}
      \sep        \ty{\tybot[\cs{o}]}
      \sep        \ty{\typlus[\cs{o}]{A}{B}}
      \sep        \ty{\tywith[\cs{o}]{A}{B}}
      \sep        \ty{\tynil[\cs{o}]}
      \sep        \ty{\tytop[\cs{o}]}
  \end{array}
\]
\textbf{Processes} ($\tm{P}$, $\tm{Q}$) in PCP are defined by the following grammar.
\[
  \begin{array}[t]{lcl}
    \tm{P}, \tm{Q}
    & \Coloneqq & \tm{\link{x}{y}}
           \sep   \tm{\res{x}{y}{P}}
           \sep   \tm{(\ppar{P}{Q})}
           \sep   \tm{\halt}
    \\   & \sep & \tm{\send{x}{y}{P}}
           \sep   \tm{\close{x}{P}}
           \sep   \tm{\recv{x}{y}{P}}
           \sep   \tm{\wait{x}{P}}
    \\   & \sep & \tm{\inl{x}{P}}
           \sep   \tm{\inr{x}{P}}
           \sep   \tm{\offer{x}{P}{Q}}
           \sep   \tm{\absurd{x}}
  \end{array}
\]
Processes are typed by sequents $\seq{P}{\ty{\Gamma}}$, which correspond to the one-sided sequents in classical linear logic. Differently from PGV, in PCP we do not need to store the greatest priority on the sequent, as, due to the absence of higher-order functions, we cannot compose processes \emph{sequentially}.

PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par} rules---corresponding to cycle and mix rules, respectively---and guarantees deadlock freedom by using priority constraints, \eg, as in \LabTirName{T-Send}. % The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics,fig:pcp-typing} in \cref{app:revisiting-PCP}:
\begin{mathpar}
  \small
  \inferrule*[lab=T-Res]{
    \seq{P}{\ty{\Gamma},\tmty{x}{A},\tmty{y}{\co{A}}}
  }{\seq{\res{x}{y}{P}}{\ty{\Gamma}}}
  \inferrule*[lab=T-Par]{
    \seq{P}{\ty{\Gamma}}
    \and
    \seq{Q}{\ty{\Delta}}
  }{\seq{\ppar{P}{Q}}{\ty{\Gamma},\ty{\Delta}}}
  \inferrule*[lab=T-Send]{
    \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
    \and
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
  }{\seq{\send{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}}}}
\end{mathpar}

@@

We prove closed progress for our updated PCP.
% The proof is in \cref{prf:thm-pcp-closed-progress}.
\begin{compacttheorems}
  \begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]%
    \label{thm:pcp-closed-progress}
    \hfill\\%newline before theorem statement
    If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
  \end{restatabletheorem}
\end{compacttheorems}
\endgroup

@@

There are two separate translations on processes. The main translation, $\tm{\cpgvM{\cdot}}$, translates processes to \emph{terms}:
\begin{align*}
  &\pcp{\tm{\cpgvM{\link{x}{y}}}}
  &&= \pgv{\tm{\link\;{\pair{x}{y}}}} \\
  &\pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
  &&= \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
  &\pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
  &&= \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
  &\pcp{\tm{\cpgvM{\halt}}}
  &&= \pgv{\tm{\unit}} \\
  &\pcp{\tm{\cpgvM{\close{x}{P}}}}
  &&= \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
  &\pcp{\tm{\cpgvM{\wait{x}{P}}}}
  &&= \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
  &\pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
  &&= \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
  &\pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
  &&= \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
  &\pcp{\tm{\cpgvM{\inl{x}{P}}}}
  &&= \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
  &\pcp{\tm{\cpgvM{\inr{x}{P}}}}
  &&= \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
  &\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
  &&= \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
  &\pcp{\tm{\cpgvM{\absurd{x}}}}
  &&= \pgv{\tm{\offeremp{x}}}
\end{align*}

Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is unsound, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which can reduce to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}. The same goes for $\nu$-binders which are inserted when translating bound send to unbound send. For instance, the process $\pcp{\tm{\send{x}{y}{P}}}$ is blocked, but its translation uses $\pgv{\tm{\new}}$ and can reduce. To address this issue, we use a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to $\tm{\cpgvM{\cdot}}$ followed by reductions using \LabTirName{E-New} and \LabTirName{E-Spawn}:
\begin{align*}
  &\pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
  &&= \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
  \\
  &\pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
  &&= \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
  \\
  &\pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
  \\
  &\pcp{\tm{\cpgvC{\inl{x}{P}}}}
  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
  \\
  &\pcp{\tm{\cpgvC{\inr{x}{P}}}}
  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
  \\
  &\pcp{\tm{\cpgvC{P}}}
  &&= \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
\end{align*}

Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$, where $\tm{\pgv{\child}}$ indicates a child thread. Translated processes do not have a main thread.
The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
% The proofs are by structural induction and can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
\begin{compacttheorems}
  \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, ${\tm{\cpgvM{\cdot}}}$]%
    \label{lem:pcp-to-pgv-terms-preservation}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
    \label{thm:pcp-to-pgv-confs-preservation}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
  \end{restatabletheorem}
  \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}%
    \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
    For any $\pcp{\tm{P}}$, either:
    \begin{itemize}
    \item $\pgv{\tm{\child\;\cpgvM{P}}=\tm{\cpgvC{P}}}$; or
    \item $\pgv{\tm{\child\;\cpgvM{P}}\cred^+\tm{\cpgvC{P}}}$, and for any $\pgv{\tm{\conf{C}}}$, if $\pgv{\tm{\child\;\cpgvM{P}}\cred\tm{\conf{C}}}$, then $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{P}}}$.
    \end{itemize}
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
    \label{thm:pcp-to-pgv-operational-correspondence-soundness}
    \hfill\\%newline before theorem statement
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
  \end{restatabletheorem}
  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
    [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
    \label{thm:pcp-to-pgv-operational-correspondence-completeness}
    \hfill\\%newline before theorem statement
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
    then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
  \end{restatabletheorem}
\end{compacttheorems}
\endgroup

\end{document}

%%% Local Variables:
%%% TeX-master: "priorities"
%%% End:

>>>>

\subsection{Revisiting Priority CP}
\begingroup
\usingnamespace{pcp}
\textbf{Types} ($\ty{A}$, $\ty{B}$) in PCP correspond to linear logic connectives annotated with priorities $\cs{o}\in\mathbb{N}$. Typing environments, duality, and the priority function $\pr(\cdot)$ are defined as expected.
\[
  \begin{array}{lcl}
    \ty{A}, \ty{B}
     & \Coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
    \sep        \ty{\typarr[\cs{o}]{A}{B}}
    \sep        \ty{\tyone[\cs{o}]}
    \sep        \ty{\tybot[\cs{o}]}
    \sep        \ty{\typlus[\cs{o}]{A}{B}}
    \sep        \ty{\tywith[\cs{o}]{A}{B}}
    \sep        \ty{\tynil[\cs{o}]}
    \sep        \ty{\tytop[\cs{o}]}
  \end{array}
\]
\textbf{Processes} ($\tm{P}$, $\tm{Q}$) in PCP are defined by the following grammar.
\[
  \begin{array}[t]{lcl}
    \tm{P}, \tm{Q}
     & \Coloneqq & \tm{\link{x}{y}}
    \sep   \tm{\res{x}{y}{P}}
    \sep   \tm{(\ppar{P}{Q})}
    \sep   \tm{\halt}
    \\   & \sep & \tm{\send{x}{y}{P}}
    \sep   \tm{\close{x}{P}}
    \sep   \tm{\recv{x}{y}{P}}
    \sep   \tm{\wait{x}{P}}
    \\   & \sep & \tm{\inl{x}{P}}
    \sep   \tm{\inr{x}{P}}
    \sep   \tm{\offer{x}{P}{Q}}
    \sep   \tm{\absurd{x}}
  \end{array}
\]
Processes are typed by sequents $\seq{P}{\ty{\Gamma}}$, which correspond to the one-sided sequents in classical linear logic. Differently from PGV, in PCP we do not need to store the greatest priority on the sequent, as, due to the absence of higher-order functions, we cannot compose processes \emph{sequentially}.

PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par} rules---corresponding to cycle and mix rules, respectively---and guarantees deadlock freedom by using priority constraints, \eg, as in \LabTirName{T-Send}.
\begin{mathpar}
  \small
  \inferrule*[lab=T-Res]{
    \seq{P}{\ty{\Gamma},\tmty{x}{A},\tmty{y}{\co{A}}}
  }{\seq{\res{x}{y}{P}}{\ty{\Gamma}}}
  \inferrule*[lab=T-Par]{
    \seq{P}{\ty{\Gamma}}
    \and
    \seq{Q}{\ty{\Delta}}
  }{\seq{\ppar{P}{Q}}{\ty{\Gamma},\ty{\Delta}}}
  \inferrule*[lab=T-Send]{
    \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
    \and
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
  }{\seq{\send{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}}}}
\end{mathpar}

@@

We prove closed progress for our updated PCP.
\begin{thm}\label{thm:pcp-closed-progress}
  %[Progress, $\pcp{\red}$]%
  If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
\end{thm}
\endgroup

@@

There are two separate translations on processes. The main translation, $\tm{\cpgvM{\cdot}}$, translates processes to \emph{terms}:
\begin{align*}
   & \pcp{\tm{\cpgvM{\link{x}{y}}}}
   &                                    & = \pgv{\tm{\link\;{\pair{x}{y}}}}                                               \\
   & \pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
   &                                    & = \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}}                                    \\
   & \pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
   &                                    & = \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}}            \\
   & \pcp{\tm{\cpgvM{\halt}}}
   &                                    & = \pgv{\tm{\unit}}                                                              \\
   & \pcp{\tm{\cpgvM{\close{x}{P}}}}
   &                                    & = \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}}                                   \\
   & \pcp{\tm{\cpgvM{\wait{x}{P}}}}
   &                                    & = \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}}                                    \\
   & \pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
   &                                    & = \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
   & \pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
   &                                    & = \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}}                              \\
   & \pcp{\tm{\cpgvM{\inl{x}{P}}}}
   &                                    & = \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}}                      \\
   & \pcp{\tm{\cpgvM{\inr{x}{P}}}}
   &                                    & = \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}}                      \\
   & \pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
   &                                    & = \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}}                              \\
   & \pcp{\tm{\cpgvM{\absurd{x}}}}
   &                                    & = \pgv{\tm{\offeremp{x}}}
\end{align*}

Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is unsound, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which can reduce to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}. The same goes for $\nu$-binders which are inserted when translating bound send to unbound send. For instance, the process $\pcp{\tm{\send{x}{y}{P}}}$ is blocked, but its translation uses $\pgv{\tm{\new}}$ and can reduce. To address this issue, we use a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to $\tm{\cpgvM{\cdot}}$ followed by reductions using \LabTirName{E-New} and \LabTirName{E-Spawn}:
\begin{align*}
   & \pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
   &                                   & = \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
  \\
   & \pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
   &                                   & = \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
  \\
   & \pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
   &                                   & = \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
  \\
   & \pcp{\tm{\cpgvC{\inl{x}{P}}}}
   &                                   & = \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
  \\
   & \pcp{\tm{\cpgvC{\inr{x}{P}}}}
   &                                   & = \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
  \\
   & \pcp{\tm{\cpgvC{P}}}
   &                                   & = \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
\end{align*}

Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$, where $\tm{\pgv{\child}}$ indicates a child thread. Translated processes do not have a main thread.
The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
\begin{lem}\label{lem:pcp-to-pgv-terms-preservation}
  %[Preservation, ${\tm{\cpgvM{\cdot}}}$]%
  If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
\end{lem}
\begin{thm}\label{thm:pcp-to-pgv-confs-preservation}
  %[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
  If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
\end{thm}
\begin{lem}\label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
  For any $\pcp{\tm{P}}$, either:
  \begin{itemize}
    \item $\pgv{\tm{\child\;\cpgvM{P}}=\tm{\cpgvC{P}}}$; or
    \item $\pgv{\tm{\child\;\cpgvM{P}}\cred^+\tm{\cpgvC{P}}}$, and for any $\pgv{\tm{\conf{C}}}$, if $\pgv{\tm{\child\;\cpgvM{P}}\cred\tm{\conf{C}}}$, then $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{P}}}$.
  \end{itemize}
\end{lem}
\begin{thm}\label{thm:pcp-to-pgv-operational-correspondence-soundness}
  %[Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
\end{thm}
\begin{thm}\label{thm:pcp-to-pgv-operational-correspondence-completeness}
  %[Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
  then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
\end{thm}
\endgroup
